---
title: "Reading Delimited Data"
format: html
toc: TRUE
editor_options: 
  chunk_output_type: inline
---
## Task 2

Load tidyverse library.
```{r}
library(tidyverse)
```

First row of data to inspect  
ADT1_YEAST 0.58 0.61 0.47 0.13 0.50 0.00 0.48 0.22 MIT

Using the width of the data column we can read in the data and assign appropiate column names.

```{r}
yeast_data <- read_fwf("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
                       fwf_widths(c(12,6,6,6,6,6,6,6,6,6),
                       col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox","vac","nuc","class")))
```

Manipulate the yeast data by starting a chain and removing the columns *seq_name* and *nuc*

```{r}
yeast_data |> 
  select(c(-"seq_name", -"nuc"))
```
Add columns for the mean and median of each numeric column at each class grouping. Note the across function allows us to the same action (make and mean and median column) to multiple columns (the numeric ones).
```{r}
yeast_data |> 
  select(c(-"seq_name", -"nuc")) |>
  group_by(class) |>
  mutate(across(where(is.numeric), list(mean = mean, md = median)))
```

